---
title: "PSY6422 Data Management and Visualisation"
author: '220225650'
date: "2023-05-04"
#output: pdf_document
#output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Background 

#Project Aims
## Research question

#Data Origins
##Why was this data chosen? 
## Codebook for both data sets
##Codebook for the rest of the project can be found...?

#Data Preperation
##Loading Packages and RENV
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
###What version of packages

##Cleaning the data - removing explanatory rows and columns that are not required
##Combining the 2017-2021 data and the 2009 to 2019 data
##Adding continent for colour choices
##Creating a rank variable and cutting off using this 
Regardless of visualisation I realised that I needed a sensible cut off value for countries. Ranks provided that.I stored as a number variable as this allowed me to trial with a few of the size of plots
## Making long format 
###2 stage process of years and then ranks
### Merging the long formats 

#Visualisation 1 - using the actual values

